---
title: "alina_trees"
author: "Alina Levine"
date: "May 15, 2019"
output: html_document
---

##regression tree with ROC

```{r setup, include=FALSE}

library(ISLR)


set.seed(1)




ctrl <- trainControl(method = "cv",
                     summaryFunction = twoClassSummary,
                     classProbs = TRUE)
set.seed(1)

tree <- train(disease~., data_train, 
                   method = "rpart",
                   tuneGrid = data.frame(cp = exp(seq(-17,2, len = 100))),
                   trControl = ctrl,
              metric = "ROC")

ggplot(tree, highlight = TRUE)
rpart.plot(tree$finalModel)

saveRDS(tree, "tree.RDS")

tree$bestTune

#highest ROC is .8003324

tree$finalModel$cptable

```

There are 15 terminal nodes

##Tree: use Kappa to tune

```{r}
set.seed(1)
ctrl_ac_kap <- trainControl(method = "cv")

set.seed(1)
tree_kappa <- train(disease~., data_train, 
                   method = "rpart",
                   tuneGrid = data.frame(cp = exp(seq(-15,-4, len = 200))),
                   trControl = ctrl_ac_kap,
                   metric = "Kappa")

ggplot(tree_kappa, highlight = TRUE)
rpart.plot(tree$finalModel)

tree_kappa$bestTune ## cp .00368

tree_kappa$results

rpart.plot(tree_kappa$finalModel)

saveRDS(tree_kappa, "tree_kappa.RDS")

#15 terminal nodes, just like ROC tuning

```


##Tree using different threholds cv

```{r}
set.seed(1)
ctrl3 <- trainControl(method = "cv", 
                      classProbs = TRUE, 
                      savePredictions = "all")


set.seed(1)
tree_kappa_thresholds <- train(disease~., data_train, 
                   method = "rpart",
                   tuneGrid = data.frame(cp = exp(seq(-15,-4, len = 200))),
                   trControl = ctrl3,
                   metric = "Kappa")

resample_stats_tree <- thresholder(tree_kappa_thresholds, 
                              threshold = seq(.05, 0.5, by = 0.05), 
                              final = TRUE)

saveRDS(resample_stats_tree, "resample_stats_tree.rds")
saveRDS(tree_kappa_thresholds, "tree_kappa_thresholds.rds")

```



###Tree Tuned by ROC Testing ROC and Kappa/Accuracy

```{r}

pred_tree_prob = predict(tree, newdata = data_test[, ,2:18], type = "prob")

pred_tree<- rep("pos", nrow(as.data.frame(pred_tree_prob)))
pred_tree[as.data.frame(pred_tree_prob)$neg>0.5] = "neg"


confusionMatrix(data = relevel(as.factor(pred_tree), ref = "pos"),
                reference = data_test[,1],
                positive = "pos")

roc.tree_test =roc(data_test$disease, as.numeric(pred_tree_prob[,1]),levels =c("neg", "pos"))

plot(roc.tree_test, legacy.axes = TRUE, print.auc = TRUE)



```


Accuracy = .9551
Kappa = .1323
sensitivity = .101695
specificity = .9897




tree using undersampling

```{r}


set.seed(1)
tree_kappa_undersample <- train(under_x_train, under_y_train, 
                   method = "rpart",
                   tuneGrid = data.frame(cp = exp(seq(-15,10, len = 200))),
                   trControl = ctrl_ac_kap,
                   metric = "Kappa")


```


#Random Forest

###Using ROC

```{r}



rf.grid <- expand.grid(mtry = 1:14,
                       splitrule = "gini",
                       min.node.size = 1:10)

set.seed(1)
rf.fit <- train(disease~., data_train, 
                method = "ranger",
                tuneGrid = rf.grid,
                metric = "ROC",
                trControl = ctrl)

saveRDS(rf.fit, "rf.fit_roc.RDS")

plot(rf.fit)

rf.fit$bestTune


```




###random forest Using Kappa

```{r}

rf.grid <- expand.grid(mtry = 1:14,
                       splitrule = "gini",
                       min.node.size = 1:10)
set.seed(1)
rf.fit_Kappa <- train(disease~., data_train, 
                method = "ranger",
                tuneGrid = rf.grid,
                metric = "Kappa",
                trControl = ctrl_ac_kap)

rf.fit_Kappa$results$Kappa

rf.fit_Kappa$bestTune

saveRDS(rf.fit_Kappa, "rf.fit_Kappa.RDS")

```

#mtry = 1, min.node.size = 1, largest kappa is 0


###random forest using kappa and threshold probability cv

```{r}

ctrl3 <- trainControl(method = "cv", classProbs = TRUE, savePredictions = "all")

rf.grid <- expand.grid(mtry = 1:14,
                       splitrule = "gini",
                       min.node.size = 1:10)
set.seed(1)
rf.fit_Kappa_threshold <- train(disease~., data_train, 
                method = "ranger",
                tuneGrid = rf.grid,
                metric = "Kappa",
                trControl = ctrl3)

saveRDS(rf.fit_Kappa_threshold, "rf_threshold_cv.rds")

resample_stats_rf <- thresholder(rf.fit_Kappa_threshold, 
                              threshold = seq(.05, 0.5, by = 0.05), 
                              final = TRUE)

saveRDS(resample_stats_rf, "resample_stats_rf.rds")
```

##random forest test set performance using roc

```{r}


pred_rf_prob = predict(rf.fit, newdata = data_test[,2:18], type = "prob")

pred_rf<- rep("pos", nrow(as.data.frame(pred_rf_prob)))

pred_rf[as.data.frame(pred_rf_prob)$neg>0.5] = "neg"

pred_rf = as.factor(c(pred_rf, "pos"))
pred_rf = pred_rf[1:1516]


confusionMatrix(data = relevel(as.factor(pred_rf), ref = "pos"),
                reference = data_test[,1],
                positive = "pos")


roc.rf_test =roc(data_test$disease, pred_rf_prob[,1],levels =c("neg", "pos"))

plot(roc.rf_test,  legacy.axes = TRUE, print.auc = TRUE)
```

test accuracy = .9611
kappa = 0
sensitivity = 0
specificity = 1
testAUC = .84



###Random forest test set performance using kappa

```{r}



pred_rf_kap = predict(rf.fit_Kappa, newdata = data_test[,2:18])


pred_rf_kap = as.factor(c(pred_rf, "pos"))
pred_rf_kap = pred_rf[1:1516]


confusionMatrix(data = relevel(as.factor(pred_rf_kap), ref = "pos"),
                reference = data_test[,1],
                positive = "pos")

```

Kappa = 0
sens: 0 
spec: 1
accuracy = .96



##boosting using ROC

```{r}

gbmB.grid <- expand.grid(n.trees = c(2000,3000,4000,5000),
                        interaction.depth = 8:10,
                        shrinkage = c(.0005, 0.001,0.003),
                        n.minobsinnode = 1)
set.seed(1)
# Binomial loss function
gbmB.fit <- train(disease~., data_train, 
                 tuneGrid = gbmB.grid,
                 trControl = ctrl,
                 method = "gbm",
                 distribution = "bernoulli",
                 metric = "ROC",
                 verbose = FALSE)

ggplot(gbmB.fit)


saveRDS(gbmB.fit, "gbmb.fit2.RDS")

boostroc = readRDS("gbmb.fit2.RDS")

```

#test set performance roc boosting

```{r}


pred_boost_prob = predict(gbmB.fit, newdata = data_test[,2:18], type = "prob")

pred_boost<- rep("pos", nrow(as.data.frame(pred_boost_prob)))

pred_boost[as.data.frame(pred_boost_prob)$neg>0.5] = "neg"


confusionMatrix(data = relevel(as.factor(pred_boost), ref = "pos"),
                reference = data_test[,1],
                positive = "pos")

roc.boost_test =roc(data_test$disease, pred_boost_prob[,1],levels =c("neg", "pos"))

plot(roc.boost_test,  legacy.axes = TRUE, print.auc = TRUE)


```

#kappa: -.0026 


#boosting using kappa

```{r}


gbmB.grid <- expand.grid(n.trees = c(2000,3000,4000,5000),
                        interaction.depth = 8:10,
                        shrinkage = c(.0005, 0.001,0.003),
                        n.minobsinnode = 1)
set.seed(1)
# Binomial loss function
gbmB.fit_kappa <- train(disease~., data_train, 
                 tuneGrid = gbmB.grid,
                 trControl = ctrl_ac_kap,
                 method = "gbm",
                 distribution = "bernoulli",
                 metric = "Kappa",
                 verbose = FALSE)
#

saveRDS(gbmB.fit_kappa, "gbmB.fit_kappa.RDS")



max_ROC_cv_boost_kappa = max(gbmB.fit$results$ROC)
tuning_param_boost_kappa = gbmB.fit$bestTune

```

max ROC is .8653252

 n.trees interaction.depth shrinkage n.minobsinnode
23    4000                10     0.001              1


#testing set

```{r}


pred_boost_kap_prob = predict(gbmB.fit_kappa, newdata = data_test[,2:18], type = "prob")

pred_boost_kap<- rep("pos", nrow(as.data.frame(pred_boost_kap_prob)))

pred_boost_kap[as.data.frame(pred_boost_kap_prob)$neg>0.5] = "neg"


confusionMatrix(data = relevel(as.factor(pred_boost_kap), ref = "pos"),
                reference = data_test[,1],
                positive = "pos")

roc.boost_kap_test =roc(data_test$disease, pred_boost_kap_prob[,1],levels =c("neg", "pos"))

plot(roc.boost_kap_test,  legacy.axes = TRUE, print.auc = TRUE)

```

AUC is .827
kappa is .0207



```{r}



gbmB.grid <- expand.grid(n.trees = c(2000,3000,4000,5000),
                        interaction.depth = 1:10,
                        shrinkage = c(.0005, 0.001,0.003),
                        n.minobsinnode = 1)
set.seed(1)
# Binomial loss function
gbmB.fit_kappa_threshold <- train(disease~., data_train, 
                 tuneGrid = gbmB.grid,
                 trControl = ctrl3,
                 method = "gbm",
                 distribution = "bernoulli",
                 metric = "Kappa",
                 verbose = FALSE)


resample_stats_gbmB <- thresholder(gbmB.fit_kappa_threshold, 
                              threshold = seq(.05, 0.5, by = 0.05), 
                              final = TRUE)

saveRDS(gbmB.fit_kappa_threshold, "gbmB.fit_kappa_threshold.rds")

saveRDS(resample_stats_gbmB, "resample_stats_gbmB.rds")


```


```{r}



gbmB.grid_expand <- expand.grid(n.trees = c(2000,3000,4000,5000),
                        interaction.depth = 8:11,
                        shrinkage = c(0.001,0.003, .005),
                        n.minobsinnode = 1)

set.seed(1)
# Binomial loss function
gbmB.fit_kappa_undersample <- train(under_x_train, under_y_train, 
                 tuneGrid = gbmB.grid_expand,
                 trControl = ctrl_ac_kap,
                 method = "gbm",
                 distribution = "bernoulli",
                 metric = "Kappa",
                 verbose = FALSE)


saveRDS(gbmB.fit_kappa_undersample, "gbmB.fit_kappa_undersample.rds")

gbmB.fit_kappa_undersample_real_expand <- train(disease~., under_data_train, 
                 tuneGrid = gbmB.grid_expand,
                 trControl = ctrl_ac_kap,
                 method = "gbm",
                 distribution = "bernoulli",
                 metric = "Kappa",
                 verbose = FALSE)


saveRDS(gbmB.fit_kappa_undersample_real_expand, "gbmB.fit_kappa_undersample_real.rds")

max(gbmB.fit_kappa_undersample_real_expand$results$Kappa)


gbmB.fit_kappa_undersample_real_expand$bestTune

#testing data




pred_boost_kap_prob = predict(gbmB.fit_kappa, newdata = data_test[,2:18], type = "prob")

pred_boost_kap<- rep("pos", nrow(as.data.frame(pred_boost_kap_prob)))

pred_boost_kap[as.data.frame(pred_boost_kap_prob)$neg>0.5] = "neg"


confusionMatrix(data = relevel(as.factor(pred_boost_kap), ref = "pos"),
                reference = data_test[,1],
                positive = "pos")



```

max kappa is .42

best tune is n.trees = 3000, interaction depth = 10, shrinkage = .003


random forest undersample

```{r}

set.seed(1)
rf.fit_undersample <- train(under_x_train, under_y_train, 
                method = "ranger",
                tuneGrid = rf.grid,
                metric = "Kappa",
                trControl = ctrl_ac_kap)

ggplot(rf.fit_undersample)


saveRDS(rf.fit_undersample, "rf.fit_undersample.RDS")

rf.fit_undersample$bestTune


```





```{r}

library(modelr) 

```

