---
title: "data_science_load_data"
author: "Alina Levine"
date: "April 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(haven)

```


Here, we will import the separate datasets from year 15 and 14 and combine for further analysis.

Our targets are participants who are 20 years older.


# Demographics variables


Age in years at screening: variable name RIDAGEYR
Gender: variable name RIAGENDR
Marital status: variable name DMDMARTL
Race/ethinicity: Race/Hispanic origin w/ NH Asian, variable name RIDRETH3
Citizenship status: variable name DMDCITZN
Education level: Education level - Adults 20+, variable name DMDEDUC2


sequence number, hispanic ethnicity, age,Ttotal household income

https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Demographics&CycleBeginYear=2015


```{r demo}
demographics_2015 = read_xpt("data/demographics.XPT") %>% 
  janitor::clean_names() %>%
  select(seqn, age = ridageyr, gender = riagendr, marital = dmdmartl, race = ridreth3, citizen = dmdcitzn, education = dmdeduc2) %>% 
  filter(age > 20)

demographics_2014 = read_xpt("data/demographics_2014.XPT") %>% 
  janitor::clean_names() %>%
  select(seqn, age = ridageyr, gender = riagendr, marital = dmdmartl, race = ridreth3, citizen = dmdcitzn, education = dmdeduc2)  %>% 
  filter(age > 20)

demographics = bind_rows(demographics_2014, demographics_2015)

```

# Income

Family monthly poverty level category, variable name: INDFMMPCSAS
https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/INQ_I.htm#INDFMMPI 

demographics = read_xpt("data/demographics.XPT") %>% janitor::clean_names()
1	Monthly poverty level index <= 1.30;2	1.30 < Monthly poverty level index <= 1.85;3	Monthly poverty level index > 1.85

```{r income}
income_2015 = read_xpt("data/income.XPT") %>% 
  janitor::clean_names() %>% 
  select(seqn, income = indfmmpc) %>% 
  mutate(income = ifelse(income %in% c(1,2,3), income, NA))

income_2014 = read_xpt("data/income_2014.XPT") %>% 
  janitor::clean_names() %>% 
  select(seqn, income = indfmmpc) %>% 
  mutate(income = ifelse(income %in% c(1,2,3), income, NA))

income = bind_rows(income_2014, income_2015)
```

#  Depression score
sum of scores for 9 items. Each item is scored as 0 (not at all) to 3 (nearly every day), which assesses the frequency of depression related symptoms.
extracted from dataset Mental Health - Depression Screener  

```{r depression}

depression_2015 = read_xpt("data/mental_health.XPT") %>% 
  janitor::clean_names()

seqn = depression_2015$seqn
depression_subset = depression_2015 %>% select(-seqn)
depression_2015 = map_dfc(depression_subset, ~ifelse(.x %in% c(0,1,2,3), .x, NA)) %>% 
  mutate(depression_score = round(rowMeans(., na.rm = TRUE), 3)) %>% 
  add_column(seqn) %>% 
  select(seqn, depression_score)

depression_2014 = read_xpt("data/mental_health_2014.XPT") %>% 
  janitor::clean_names()
seqn = depression_2014$seqn
depression_subset = depression_2014 %>% select(-seqn)
depression_2014 = map_dfc(depression_subset, ~ifelse(.x %in% c(0,1,2,3), .x, NA)) %>% 
  mutate(depression_score = round(rowMeans(., na.rm = TRUE), 3)) %>% 
  add_column(seqn) %>% 
  select(seqn, depression_score)

depression = bind_rows(depression_2014, depression_2015)
```

#blood pressure and cholesterol
ever told you have high blood pressure, doctor ever told you have high cholesterol

https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/BPQ_I.htm

```{r}

bp_chol_2015 = read_xpt("data/bp_chol.xpt") %>% janitor::clean_names() %>%
  select(seqn, highbp = bpq020, highchol = bpq080) %>%
  mutate(highbp = ifelse(highbp == 9, NA, highbp),
         highchol = ifelse(! (highchol %in% c(1,2)), NA, highchol))

bp_chol_2014 = read_xpt("data/bp_chol_2014.xpt") %>% janitor::clean_names() %>%
  select(seqn, highbp = bpq020, highchol = bpq080) %>%
  mutate(highbp = ifelse(highbp == 9, NA, highbp),
         highchol = ifelse(! (highchol %in% c(1,2)), NA, highchol))

bp_chol = bind_rows(bp_chol_2014, bp_chol_2015)
```


#smoking
smoked at least 100 cigarettes in life

```{r}

smoking_2015 = read_xpt("data/smoking.xpt") %>% janitor::clean_names() %>%
  select(seqn, smoked = smq020) %>%
  mutate(smoked = ifelse(!(smoked %in% c(1,2)), NA, smoked))
  
smoking_2014 = read_xpt("data/smoking_2014.xpt") %>% janitor::clean_names() %>%
  select(seqn, smoked = smq020) %>%
  mutate(smoked = ifelse(!(smoked %in% c(1,2)), NA, smoked))
 
smoking = bind_rows(smoking_2014, smoking_2015)                                   
```



#alcohol

had at least 12 drinks of alcohol/1 year (ever a moderate drinker), ever had 4/5 drinks every day

```{r}

alcohol_2015 = read_xpt("data/alcohol.xpt") %>% janitor::clean_names() %>%
  select(seqn,drinker = alq101, heavy_drinker = alq151 ) %>%
  mutate(heavy_drinker = ifelse(!(heavy_drinker %in% c(1,2)), NA, heavy_drinker),
         drinker = ifelse(!(drinker %in% c(1,2)), NA, drinker))

alcohol_2014 = read_xpt("data/alcohol_2014.xpt") %>% janitor::clean_names() %>%
  select(seqn,drinker = alq101, heavy_drinker = alq151 ) %>%
  mutate(heavy_drinker = ifelse(!(heavy_drinker %in% c(1,2)), NA, heavy_drinker),
         drinker = ifelse(!(drinker %in% c(1,2)), NA, drinker))
alcohol = bind_rows(alcohol_2014, alcohol_2015)
```


# Diet Behavior and Nutrition

```{r}

diet_behavior_2015 =
  read_xpt("data/diet.xpt") %>% 
  select(seqn = SEQN, 
         fast = DBD900, 
         ready = DBD905, 
         forzen = DBD910) %>% 
  filter(!fast %in% c(5555, 7777, 9999),
         !ready %in% c(6666, 7777, 9999),
         !forzen %in% c(6666, 7777, 9999))


# proportion of missing value
sum(is.na(diet_behavior_2015$fast))/length(diet_behavior_2015$seqn)
sum(is.na(diet_behavior_2015$ready))/length(diet_behavior_2015$seqn)
sum(is.na(diet_behavior_2015$forzen))/length(diet_behavior_2015$seqn)
```

```{r}

diet_behavior_2014 =
  read_xpt("data/diet_2014.xpt") %>% 
  select(seqn = SEQN, 
         fast = DBD900, 
         ready = DBD905, 
         forzen = DBD910) %>% 
  filter(!fast %in% c(5555, 7777, 9999),
         !ready %in% c(6666, 7777, 9999),
         !forzen %in% c(6666, 7777, 9999))


diet_behavior = rbind(diet_behavior_2014, diet_behavior_2015)

# proportion of missing value
sum(is.na(diet_behavior$fast))/length(diet_behavior$seqn)
sum(is.na(diet_behavior$ready))/length(diet_behavior$seqn)
sum(is.na(diet_behavior$forzen))/length(diet_behavior$seqn)
```


# Body Measure

```{r}
body_measures_2015 = 
  read_xpt("data/body_measures.xpt") %>% 
  select(seqn = SEQN, 
         bmi = BMXBMI)

# proportion of missing value
sum(is.na(body_measures_2015$bmi))/length(body_measures_2015$seqn)
```

```{r}
body_measures_2014 = 
  read_xpt("data/body_measures_2014.xpt") %>% 
  select(seqn = SEQN, 
         bmi = BMXBMI)

body_measures = rbind(body_measures_2014, body_measures_2015)
```

# Diabetes

```{r}

diabetes_2015 =
  read_xpt("data/diabetes.xpt") %>% 
  select(seqn = SEQN, 
         diabete = DIQ010) %>% 
  mutate(diabete = ifelse(!(diabete %in% c(1,2,3)), NA, diabete))


diabetes_2014 =
  read_xpt("data/diabetes_2014.xpt") %>% 
  select(seqn = SEQN, 
         diabete = DIQ010) %>% 
  mutate(diabete = ifelse(!(diabete %in% c(1,2,3)), NA, diabete))

diabetes = rbind(diabetes_2014, diabetes_2015)

# proportion of missing value
sum(is.na(diabetes$diabete))/length(diabetes$seqn)
```

# Medical Conditions

```{r}

medical_cond_2015 =
  read_xpt("data/medical_cond.xpt") %>% 
  select(seqn = SEQN, 
         overweight = MCQ080, 
         gout = MCQ160N) %>% 
  mutate(overweight = ifelse(!(overweight %in% c(1,2)), NA, overweight),
         gout = ifelse(!(gout %in% c(1,2)), NA, gout))

medical_cond_2014 =
  read_xpt("data/medical_cond_2014.xpt") %>% 
  select(seqn = SEQN, 
         overweight = MCQ080, 
         gout = MCQ160N) %>% 
  mutate(overweight = ifelse(!(overweight %in% c(1,2)), NA, overweight),
         gout = ifelse(!(gout %in% c(1,2)), NA, gout))

medical_cond = rbind(medical_cond_2014, medical_cond_2015)

# proportion of missing value
sum(is.na(medical_cond$overweight))/length(medical_cond$seqn)
sum(is.na(medical_cond$gout))/length(medical_cond$seqn)
```



# Response

```{r}
coronary_2015 = 
  read_xpt("data/medical_cond.xpt") %>% 
  select(seqn = SEQN, 
         disease = MCQ160C) %>% 
  mutate(disease = ifelse(!(disease %in% c(1,2)), NA, disease))


coronary_2014 = 
  read_xpt("data/medical_cond_2014.xpt") %>% 
  select(seqn = SEQN, 
         disease = MCQ160C) %>% 
  mutate(disease = ifelse(!(disease %in% c(1,2)), NA, disease))

coronary = rbind(coronary_2014, coronary_2015)

sum(is.na(coronary$disease))/length(coronary$seqn)
```



#variables we might need to remove: fast food, heavy drinker(but this is important one)



```{r}

full_data = plyr::join_all(list(demographics, income, depression, diabetes, smoking, diet_behavior, alcohol, medical_cond, coronary, body_measures), by = "seqn", type = "inner") %>% 
  filter(!is.na(disease)) %>% 
  dplyr::select(-seqn) %>% 
  na.omit

skimr::skim(full_data)

# 6115 records, 18 variables 
```



#Recoding
now we are changing yes to 1 and no to 0. changing all binary to 0 1 and all categorical to factors

```{r}
  
full_data_binary_subset = full_data %>% 
  select(c(gender, citizen, smoked, drinker, heavy_drinker,overweight,gout, disease))

full_data_non_binary = full_data %>%
  select(-c(gender, citizen,smoked, drinker, heavy_drinker,overweight,gout, disease))

full_data = cbind(map_dfc(full_data_binary_subset, ~as.factor(ifelse(.x==2, 0, .x))), full_data_non_binary) %>%
  mutate(marital = as.factor(marital),
         race = as.factor(race),
         income = as.factor(income),
         education = as.factor(education),
         diabete = as.factor(diabete)) %>% 
  mutate(disease = recode(disease, "0" = "neg", "1" = "pos")) %>% 
  select(disease, everything())

# write.csv(full_data, "data/full_data.csv", row.names = FALSE)
```

#split training and test dataset


```{r}

library(caret)
set.seed(1)
rowTrain <- createDataPartition(y = full_data$disease,
                                p = 0.75,
                                list = FALSE)

data_train = full_data[rowTrain,]
data_test = full_data[-rowTrain,]


```
